---
title: "Store Data Analysis"
author: "Akis Gazepids"
date: "15/1/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Wrangling

Σκοπός της εργασίας είναι η διερεύνηση των δεδομένων λιανεμπορίου (Exploratory Data Analysis). Ειδικότερα θα πρέπει να διαχειριστείτε τα διαθέσιμα δεδομένα ώστε να παραχθούν γραφικές παραστάσεις συνοδευόμενες από σχολιασμό θα πρέπει να απαντούν στα ερωτήματα.

Categories.csv: Τα προϊόντα (ean) είναι οργανωμένα σε προϊοντικές κατηγορίες 4 επιπέδων. Απεικονίζει μια δενδροειδή δομή οργάνωση των προϊόντων. Κάθε μια γραμμή είναι ένα προϊόν και τα δεδομένα που ακολουθούν είναι το μονοπάτι από μια γενική προϊοντική κατηγορία σε μια ειδίκευση. 

Sales_464.csv (και Sales_493.csv): Περιέχει τις πωλήσεις που έχουν γίνει σε κάθε κατάστημα για διάστημα περίπου 12 μηνών. 

### Insert Data

```{r}
library(ggplot2)
library(dplyr)
library("data.table")
options(knitr.duplicate.label = "allow")
categories_df = read.csv("C:\\Users\\Intel PC G5900\\Desktop\\Data\\Categories.csv",row.names=NULL, encoding="UTF-8",sep=";", stringsAsFactors=FALSE,header = TRUE)
sales_464_df = read.csv("C:\\Users\\Intel PC G5900\\Desktop\\Data\\Sales_464.csv",sep=";")
sales_493_df = read.csv("C:\\Users\\Intel PC G5900\\Desktop\\Data\\Sales_493.csv",sep=";")

```

### Print the dimensions of the three dataframes
```{r}
print(dim(categories_df))
print(dim(sales_464_df))
print(dim(sales_493_df))
```

### Print top 10 rows of each dataframe
```{r}
print(head(categories_df , 5))
print(head(sales_464_df , 5))
print(head(sales_493_df , 5))
```

### Print the column names of each dataframe
```{r}
print(colnames(categories_df))
print(colnames(sales_464_df))
print(colnames(sales_493_df))
```

### Print the types of each dataframe
```{r}
print(sapply(categories_df, typeof))
print(sapply(sales_464_df, typeof))
print(sapply(sales_493_df, typeof))
```

### Group by ean and sum quantity and price
```{r}
agg_tbl <- sales_464_df %>% group_by(ean) %>% 
  summarise(total_price=sum(price),
            total_quantity = sum(quantity),.groups = 'drop')

sales_464_grouped_df <- agg_tbl %>% as.data.frame()
```

### Sort by quantity in descending order and keep the top 20% of quantity sales
```{r}
sales_464_sortedbyQyantity_df = sales_464_grouped_df[order(-sales_464_grouped_df$total_quantity),]
sales_464_20_perc = head(sales_464_sortedbyQyantity_df, round(dim(sales_464_sortedbyQyantity_df)[1] *0.2))
```

### Keep from the initial dataframe only the codes of the ean column after keeping only the top 20%
```{r}
list= sales_464_20_perc$ean
filtered_sales_464_df = sales_464_df[sales_464_df$ean %in% list ,]
```

### Group by date and sum prices in the filtered dataframe and sort by date in ascending order
```{r}
agg_tbl <- filtered_sales_464_df %>% group_by(date_trns) %>% 
  summarise(total_price=sum(price),.groups = 'drop')

sales_464_grouped_df <- agg_tbl %>% as.data.frame()
sales_464_sorted_df = sales_464_grouped_df[order(sales_464_grouped_df$date_trns),]
```

### Convert date_trns from character to date
```{r}
sales_464_sorted_df$date_trns <- as.Date(sales_464_sorted_df$date_trns, "%Y-%m-%d")
```

## Create and print scatterplot
1.	Sample of 50 points in a scatterplot of the daily revenue of top 20% of the products in terms of quantity count.
```{r scatterplot, echo=FALSE}
ggplot(data = head(sales_464_sorted_df, 50), aes(date_trns, total_price)) +
      geom_point() +
      theme(axis.text.x = element_text(angle = 75, vjust = 0.5, hjust=1))+
      ylim(0, 20000)
```

# Convert date_trns from character to date to inital dataframe
```{r}
sales_464_df$date_trns = as.Date(sales_464_df$date_trns, "%Y-%m-%d")
```
#Create a column for month and week number of the date 
```{r}
sales_464_df$month =format(sales_464_df$date_trns, "%b")
sales_464_df$weekNum =format(sales_464_df$date_trns, "%V")
```

# Group by month in the dataframe and sum prices and quantity
```{r}
agg_tbl <- sales_464_df %>% group_by(month) %>% 
  summarise(total_price=sum(price),
            total_quantity = sum(quantity),.groups = 'drop')

sales_464_grouped_month_df <- agg_tbl %>% as.data.frame()
```
# Create and print barplot
1.	Monthly revenue of store 464 in a barplot.
```{r barplot, echo=FALSE, fig.keep='all'}
ggplot(sales_464_grouped_month_df, aes(x=month, y=total_price)) + 
  geom_bar(stat = "summary", fun = "median")+
  scale_y_continuous(name="Monthly revenue of store", labels = scales::comma)
```
